---
title: "BLS Data Exploration"
author: "Wat Hughes"
date: "August 1, 2015"
output: html_document
---

## Introduction

This shows the str, summary, head, and tail of each of the BLS datafiles for the DDL3 project. It uses a local cache of the data rather than loading > 6 GB each time it is run. Scroll down past the data load messages to see the sections for each kind of exploratory output.

```{r}
# Modify these to fit your environment:
DDLRoot = 'e:/wat/misc/ddl'
ProjectDir = paste0(DDLRoot,'/03-the-redline')
DataDir = paste0(DDLRoot,'/Data')
OrigDataDir = paste0(DataDir,'/BLSOrig')
FilesListURL ='http://download.bls.gov/pub/time.series/cs'
MaxRowToRead = 1000000 # data max nearly 50M but that takes a while to load
HeadTailN = 10 # Head and Tail sections dipslay this many rows from each dataset
```

## Data Load Messages

```{r, echo=FALSE}
library(data.table)

DFs = list()
for(FileName in dir(OrigDataDir))
{
    print(FileName)
    if (FileName %in% c('cs.contacts','cs.txt','cs.data.0.Current')) next # These do not contain tabular data or duplicate other files.

    FilePath = paste(OrigDataDir, FileName, sep='/')
    if (file.size(FilePath) > 999999)
    {
        DF = fread(FilePath,nrow=MaxRowToRead)
    }
    else
    {
        DF = read.table(FilePath,header=T,nrow=MaxRowToRead,sep='\t',row.names=NULL)
    }
    # I am not sure why but even though some files load correctly, some start with
    # a column "row.names" and right shift all the column names, with the last column
    # name being assigned to a made up column of all NAs. I am sure this is related to
    # the trailing tabs on every line except the header. Still, why inconstint loads?
    cn = colnames(DF)
    if (cn[1] == 'row.names')
    {
        DF = DF[,1:ncol(DF)-1]
        colnames(DF) = cn[2:length(cn)]
    }
    DFs[[length(DFs)+1]] = list(FileName, DF)
}
```

## Structure

```{r, echo=FALSE}
# This is used for the print side effect of cat. Junk is ignored.
junk = sapply(DFs, function(lDF) {(cat('\n',lDF[[1]],' ')); str(lDF[[2]])})
```

## Summary

```{r, echo=FALSE}
# This is used for the print side effect of cat. Junk is ignored.
junk = sapply(DFs, function(lDF) {cat('\n',lDF[[1]],'\n'); print(summary(lDF[[2]]))})
```

## Head

```{r, echo=FALSE}
junk = sapply(DFs, function(lDF) {cat('\n',lDF[[1]],'\n'); print(head(lDF[[2]],n=HeadTailN))})
```

## Tail

(For those datasets not completely displayed in the Head section.)

```{r, echo=FALSE}
junk = sapply(DFs, function(lDF) if (nrow(lDF[[2]]) > HeadTailN) {cat('\n',lDF[[1]],'\n'); print(tail(lDF[[2]],n=HeadTailN))})
```
